Optimal management of a stochastically varying population when policy adjustment is costly

### 2015-05-13 manuscripts/manuscript.tex
tex file

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James N. Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul R. Armsworth 
    affiliation: UTKEEB

address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
  - code: UCD
    address: Department of Environmental Science and Policy, University of California, Davis
  - code: melbourne
    address: School of Botany, University of Melbourne, Australia 
  - code: UTKEEB
    address: Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville
  - code: UTKEcon
    address: Department of Economics, University of Tennessee, Knoxville

abstract: |
          Ecological systems are dynamic and policies to manage them need to
          respond to that variation. However, policy adjustments will sometimes
          be costly, which means that fine-tuning a policy to track variability in
          the environment very tightly will only sometimes be worthwhile. We use a
          classic fisheries management question -- how to manage a stochastically
          varying population using annually varying quotas in order to maximize
          profit  -- to examine how costs of policy adjustment change optimal
          management recommendations. Costs of policy adjustment (here changes in
          fishing quotas through time) could take different forms. For example,
          these costs may respond to the size of the change being implemented, or
          there could be a fixed cost any time a quota change is made. We show
          how different forms of policy costs have contrasting implications for
          optimal policies. Though it is frequently assumed that costs to adjusting
          policies will dampen variation in the policy, we show that certain cost
          structures can actually increase variation through time. We further show
          that failing to account for adjustment costs has a consistently worse
          economic impact than would assuming these costs are present when they are not.

keywords: [optimal control, fisheries, ecological management, adjustment costs]

layout: review,12pt,author-year,titlepage
linenumbers: true
bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle
geometry: [verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,l
margin=2.54cm,rmargin=2.54cm]

output:
  pdf_document:
    template: components/elsarticle.latex
    fig_caption: true
    keep_tex: true
---

```{r configure, include=FALSE, cache=FALSE}
library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

library("rmarkdown")
library("pdgControl")
library("reshape2")
library("plyr")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
library("ggthemes")
library("snowfall")
library("tidyr")
library("dplyr")
library("gtable")
library("gridExtra")
library("parallel")
options(digits=2)

theme_set(theme_tufte(base_size=12) + theme(axis.line = element_line()))
```

```{r parameters, cache=FALSE}
parallel = TRUE
ncpu = parallel::detectCores()
price = 10
c0 = 30
profit <- profit_harvest(price = price, c0 = c0, c1 = 0)
#c2 <- exp(seq(0, log(41), length.out = 40))-1
c2 <- seq(0, 40, length.out=100)
reduction <- 0.25 
seed <- 123                 # Random seed (replicable results)
delta <- 0.05               # economic discounting rate
OptTime <- 50               # stopping time
gridsize <- 50              # grid size for fish stock and harvest rate (discretized population)
sigma_g <- 0.2              # Noise in population growth
reward <- 0                 # bonus for satisfying the boundary condition
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() 1         # No measurement noise, 
z_i <- function() 1         # No implemenation noise
f <- BevHolt                # Select the state equation
pars <- c(1.5, 0.05)        # parameters for the state equation
K <- (pars[1] - 1)/pars[2]  # Carrying capacity (for reference 
xT <- 0                     # boundary conditions
x0 <- K
x_grid <- seq(0.01, 1.2 * K, length = gridsize)  
h_grid <- seq(0.01, 0.8 * K, length = gridsize)  
```

```{r analysis, include=FALSE}
source("components/analysis.R")
```

Introduction
============

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space [@Durrett1994]. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2010]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to more closely reflect environmental variation
should be traded off against potential costs associated with the more
responsive approach to management this would require.

```{r Figure_1, fig.cap="Estimated stock abundances (top panels) show much greater variation than the corresponding harvest policies (bottom).  Left panels: ICCAT Stock assessments and Total Allowable Catch (TAC, before adjustments for national underages and overages) for bluefin tuna, 1982 - 2007. Right panels: SEDAR stock assessments and annual quotas for south Atlantic snowy grouper, 1994-2012.", fig.width=6, fig.height=4, dev=c("pdf", "svg", "png"), cache=FALSE}
tuna <- read.csv("components/data/tuna.csv")
grouper <- read.csv("components/data/grouper.csv")

ex <- rbind(cbind(tuna, species="tuna"), cbind(grouper, species="snowy grouper"))
ex_long <- gather(ex, variable, value, stock, quota)

ggplot(ex_long, aes(year,value)) + 
  geom_line() + geom_point() + 
  facet_wrap(variable~species, scales="free") + 
  ylab("value (mt)") + theme_bw()
```

To illustrate this tradeoff, we use an example from fisheries
management. Figure 1, top-left panel, shows a time series for the
estimated population size, represented as spawning stock biomass, of
west Atlantic bluefin tuna (_Thunnus thynnus_, henceforth bluefin) from
a recent stock assessment [@ICCAT2009a].  The bottom-left panel 
shows the catch quota (Total Allowable Catch or TAC, before being adjusted
for overages or underages in the catch of each country in the previous year)
for the stock as set by the relevant
management agency [@ICCAT2009a]. Despite the estimated population size
declining by `r 100 * (1 - min(tuna$stock) / max(tuna$stock))`% between
1982 and 1991, the quota was not changed during
this period. Instead, the quota only changed occasionally and in between
times was left unaltered. Fishery management decisions regarding this
species can be highly contentious (@Safina1998, @Sissenwine1998,
@Porch2005). Moreover, the stock is fished by fleets from many nations
with quotas being set by a multilaterial management agency through a
process of negotiation. As such, we might reasonably anticipate that
for this species there could be substantial transaction costs involved
in reaching agreement over any quota change, or constraints on the frequency
with which the those changes could occur.  Either could contribute to
the observed quota stability.

Such patterns of variable stocks managed
under far less variable quotas are extremely common; for instance, 
in the management of south Atlantic snowy grouper (_Epinephelus niveatus_) 
shown in the righthand panels [@SEDAR2013]. Though the stock assessment
shows similarly frequent variation between 1994 and 2012, the quota has
only been adjusted twice during this period. In contrast to the adjustment
in the bluefin quotas, both adjustments in snowy grouper quotas were phased
in through steps over several years: 1994-96 and 2005-08.
More generally, reviews by @Biais1995 and
@Patterson2007a document many cases where changes in catch
quotas that a management agency set were more modest than changes that
would be recommended just by considering variations in stock abundance.

We use a classic fisheries management question to examine how accounting
for costs of policy adjustment can change optimal policies (see also
@Ludwig1980, @Feichtinger1994, @Wirl1999). We focus on how
harvest quotas for a stochastically varying fish population can be
chosen to maximize the net present value of a fishery. Our formulation
and solution method largely follow Reed's classic treatment on this
question, a treatment repeated widely in bioeconomic textbooks. We 
note that while later work has extended this treatment to deal with 
a variety of other issues (e.g. @Sethi2005; @Singh2006; @McGough2009),
we start from the classical model for simplicity of presentation and analysis.

With his formulation, @Reed1979 showed that a constant escapement policy could be
optimal under certain conditions. Such a policy involves choosing annual
quotas that are perfectly responsive to recruitment variation in a fish
stock. In poor recruitment years, the quota is set to zero and no fishing
is allowed. Any time a recruitment pulse exceeds the escapement threshold, a quota is
set that allows the fishery to exactly compensate through harvesting,
thereby maintaining the optimal escapement level. However, in that
analysis, Reed did not account for any costs of policy adjustment, which
for such a responsive management strategy potentially could be large.

@@

Methods
=======

Fish population dynamics (state equation)
-----------------------------------------

@@

where $N_t$ the stock size, $h_t$ the harvested level, 
$Z_t$ gives the stochastic shocks, which we
assume are log-normally distributed and $A$ and $B$ are positive
constants.

We assume managers set an annual quota for harvesting $h_t$ (the control
variable) after observing the stock size that year $N_t$ (the state
variable), but while still being uncertain about future environmental
conditions and stock sizes. Through time, this gives a time path of
management actions ${\bf h}=(h_1,h_2,\dots)$ that depend on the stock
sizes that were observed (a state dependent control rule).

@@

where $\mathbf{E}$ is the expectation operator, $\delta>0$ is the
discount rate and $\Pi_0$ is the net revenue from operating the fishery
in a given year. In this base case, we assume the annual dockside (that is,
before accounting for any adjustment costs that may be incurred) net revenue
from fishing is

@@

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t = q E_t N_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. \eqref{eq:annual_netrevenue}, $p$ is the price per unit harvest and
$c_0$ the cost of fishing per unit effort and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$.

To simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the equilibrium biomass for the equivalent
deterministic model without harvesting is `r (pars[1]-1)/pars[2]`; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn \eqref{eq:state_equation}). To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.
In the supplement we illustrate that the qualitative patterns observed
here are not sensitive to the specific choices of these parameter values.

Taken together this objective function \eqref{eq:objective} and the state
equation \eqref{eq:state_equation} define a stochastic dynamic programming
problem that we solve using backwards recursion via Bellman's equation.
We denote the resulting state-dependent, optimal control as $\mathbf{h}_0^{\ast}$.
We solve this problem on a finite time horizon of $T =$ `r OptTime` using
value iteration [@Mangel1988; @Clark2000].

R code for implementing the dynamic programming algorithm in this context is
provided as a supplementary R package to the paper, along with scripts for replicating 
the analyses presented here. To address concerns about computational reproducibility [@Boettiger2015],
a Docker container image of the software environment is also provided.

<!-- Cite & link to Zenodo DOI here(?)-->

Costs of policy adjustment 
--------------------------

We compare this base case to three alternative problem formulations,
each reflecting different plausible functional forms that costs of
policy adjustment could take. In each, we assume managers can adjust the
quota set in the fishery $h_t$ in a given year and that any policy
adjustment costs are associated with changes to this control variable.
In each case, we assume there is no cost in initially setting the harvest
policy at time 0.

First we assume that policy adjustment costs are directly proportional
to the magnitude of the change in policy being proposed, such that
larger changes to annual harvesting quotas incur greater policy
adjustment penalties. Specifically, we replace $\Pi_0$ in the $NPV_0$
equation with

@@

where indicator function $\mathbf{I}(h_t, h_{t-1})$ takes the value 1 if
$h_t=h_{t-1}$ and zero otherwise. For each $\Pi_i$, we can then define a
new objective function $NPV_i$ similar to that in Eqn. \eqref{eq:objective}.
We note that the fixed cost has conceptual analogues to the set-up cost of
@Reed1974 and @Spulber1982 but here the fee is associated with a change
in harvesting ($h_t\neq h_{t-1}$) and not with harvesting *per se*
($h_t>0$) as in those studies.

Taken together with the state equation Eqn. \eqref{eq:state_equation}, each
of these new objective functions defines a different stochastic dynamic
programming problem. Again, we solve them numerically using backwards
recursion. To include costs of policy adjustment, we expand the state
space to include both the current stock size and the management action
taken on the previous time step $(N_t,h_{t-1})$. For each new objective
function $\max NPV_i$, we denote the corresponding optimal control
policy as the vector $\mathbf{h}_i^*$.

Note that this problem is much larger computationally than the classic 
formulation of this stochastic dynamic programming problem for a single stock.
Whereas the classic problem considers each of $Q$ possible discrete quotas
for $S$ possible stock values at each time $t$, for a search space of size $S \times Q$
per timestep;  this formulation must also consider all possible values of quota
in the _previous_ time step: since the costs that will follow depend on whether
and by how much the harvest policy will change. This creates a total of
$S \times Q \times Q$ configurations.

In the Supplementary Material, we compare our results with more
conventional fisheries economic formulation in which additional costs are
applied to the control variables themselves, as opposed to adjustments to
the controls: $\Pi_4 (N_t,h_t) = p h_t -  c_0 E_t-c_4E_t^2$.  Additional
costs of this form tend to have a smoothing effect on optimal quotas,
but also change the long-term average stock size or quota size that
is optimal.

Comparing apples to apples
--------------------------

Each policy adjustment cost function is characterized in terms of a
cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect
of each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal
fishery $NPV$.  Fortunately, the optimal control framework provides a
completely natural way to make this comparison: in terms of the _economic
value_ (measured by $NPV$) of the stock in each case.  We will calibrate
our realized economic value under each policy relative to economic value
of the stock when adjustments to the policy are free ($NPV_0$).

Figure 2 shows the calibration graphically. Each curve plots the change in
maximum expected $NPV$ for a given cost structure as the $c_i$ coefficient
is increased. In each case, the figure shows maximum expected
NPV with policy adjustment costs as a proportion of the maximum expected
NPV available in the basic problem $NPV_0({\bf h_0^*})$ without policy
adjustment costs.

```{r Figure_2, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost, $\\tfrac{NPV_0 - NPV_i}{NPV_0}$.  The horizontal line indicates a value of the stock that is reduced by 25% from the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=5, fig.height=3, cache = fig.cache, dev=c("pdf", "svg", "png")}

relabel <- c(L1 = substitute(paste(Pi[1])),
             L2 = substitute(paste(Pi[2])),
             fixed = substitute(paste(Pi[3])))

ggplot(fees, aes(c2, (npv0-value)/npv0, lty=variable, color=variable)) + 
  geom_line() + 
  geom_hline(aes(yintercept=reduction), linetype=4) + 
  xlab("Penalty coefficient") +
  ylab("Reduction in net present value") + 
  scale_linetype_discrete(labels=relabel, breaks = c('L1', 'L2', 'fixed'), name = 'cost model') + 
  scale_color_discrete(labels=relabel, breaks = c('L1', 'L2', 'fixed'), name = 'cost model') + 
  annotate("text", mean(c2), reduction+.02, label = "25% reduction in NPV")

write.csv(fees, "components/data/figure2.csv")

```

To compare the impact of penalty functions on optimal management across
the different functional forms, we select penalty cost coefficients
that induce the same reduction in maximum expected NPV. For example,
the dashed vertical line in Figure 2 maps the needed cost coefficient
$c_i$, for each penalty function, such that the fishery is worth `r 100*(1-reduction)`%
of its unconstrained value when optimally managed. To demonstrate that the results
shown here are independent of this choice of calibrating to a `r 100*(1-reduction)`%
reduction, we conduct the same analysis with a 10% and 30% reduction rate as well.
The corresponding figures can be found in the Supplemental Materials.

Results
=======

Effect of policy adjustment costs on optimal quotas and stock sizes 
-------------------------------------------------------------------

>>>>

\documentclass[review,12pt,author-year,titlepage]{components/elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add 
  \linenumbers % turns line numbering on 
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,l margin=2.54cm,rmargin=2.54cm]{geometry}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Optimal management of a stochastically varying population when policy adjustment is costly},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\usepackage[nomarkers]{endfloat}

\begin{document}
\begin{frontmatter}

  \title{Optimal management of a stochastically varying population when policy
adjustment is costly}
    \author[cstar]{Carl Boettiger\corref{c1}}
   \ead{cboettig (at) gmail.com} 
   \cortext[c1]{Corresponding author}
    \author[melbourne]{Michael Bode}

    \author[UCD]{James N. Sanchirico}

    \author[UTKEcon]{Jacob LaRiviere}

    \author[UCD]{Alan Hastings}

    \author[UTKEEB]{Paul R. Armsworth}

      \address[cstar]{Center for Stock Assessment Research, 110 Shaffer Rd, Santa Cruz, CA
95050, USA}    
    \address[UCD]{Department of Environmental Science and Policy, University of
California, Davis}    
    \address[melbourne]{School of Botany, University of Melbourne, Australia}    
    \address[UTKEEB]{Department of Ecology and Evolutionary Biology, University of Tennessee,
Knoxville}    
    \address[UTKEcon]{Department of Economics, University of Tennessee, Knoxville}

  \begin{abstract}
  Ecological systems are dynamic and policies to manage them need to
  respond to that variation. However, policy adjustments will sometimes be
  costly, which means that fine-tuning a policy to track variability in
  the environment very tightly will only sometimes be worthwhile. We use a
  classic fisheries management question -- how to manage a stochastically
  varying population using annually varying quotas in order to maximize
  profit -- to examine how costs of policy adjustment change optimal
  management recommendations. Costs of policy adjustment (here changes in
  fishing quotas through time) could take different forms. For example,
  these costs may respond to the size of the change being implemented, or
  there could be a fixed cost any time a quota change is made. We show how
  different forms of policy costs have contrasting implications for
  optimal policies. Though it is frequently assumed that costs to
  adjusting policies will dampen variation in the policy, we show that
  certain cost structures can actually increase variation through time. We
  further show that failing to account for adjustment costs has a
  consistently worse economic impact than would assuming these costs are
  present when they are not.
  \end{abstract}
   \begin{keyword} optimal control \sep fisheries \sep ecological management \sep adjustment costs \sep \end{keyword}
 \end{frontmatter}

\begin{flushleft}
\section{Introduction}\label{introduction}

Ecosystems are dynamic and exhibit rich patterns of variability in both
time and space (Durrett and Levin 1994). In designing management
policies for ecosystems, managers need to decide how much of that
variation to respond to. Managers could try to track variations in
ecosystem dynamics very closely by setting policies that are extremely
responsive to the environment. Many theoretical studies that seek to
identify optimal policies for exploited populations and communities
adopt this approach and largely ignore the challenges that would be
involved in implementing such recommendations (e.g., Reed 1979, Neubert
2003, Sethi et al. 2005, Halpern et al. 2011). However, the policy
process can often be much more sluggish to respond to variations in
ecosystem dynamics (Walters 1978, Armsworth et al. 2010). Moreover,
stakeholders impacted by ecosystem management may prefer some stability
and not want to deal with continually changing management
recommendations (Biais 1995, Armsworth and Roughgarden 2003, Patterson
2007, Patterson and Resimont 2007, Sanchirico et al. 2008). In other
words, whatever gains are available from fine-tuning a policy
prescription to more closely reflect environmental variation should be
traded off against potential costs associated with the more responsive
approach to management this would require.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-Figure_1-1.pdf}
\caption{Estimated stock abundances (top panels) show much greater
variation than the corresponding harvest policies (bottom). Left panels:
ICCAT Stock assessments and Total Allowable Catch (TAC, before
adjustments for national underages and overages) for bluefin tuna, 1982
- 2007. Right panels: SEDAR stock assessments and annual quotas for
south Atlantic snowy grouper, 1994-2012.}
\end{figure}

To illustrate this tradeoff, we use an example from fisheries
management. Figure 1, top-left panel, shows a time series for the
estimated population size, represented as spawning stock biomass, of
west Atlantic bluefin tuna (\emph{Thunnus thynnus}, henceforth bluefin)
from a recent stock assessment (International Commission for the
Conservation of Atlantic Tunas (ICCAT) 2009). The bottom-left panel
shows the catch quota (Total Allowable Catch or TAC, before being
adjusted for overages or underages in the catch of each country in the
previous year) for the stock as set by the relevant management agency
(International Commission for the Conservation of Atlantic Tunas (ICCAT)
2009). Despite the estimated population size declining by 53.3\% between
1982 and 1991, the quota was not changed during this period. Instead,
the quota only changed occasionally and in between times was left
unaltered. Fishery management decisions regarding this species can be
highly contentious (Safina (1998), Sissenwine et al. (1998), Porch
(2005)). Moreover, the stock is fished by fleets from many nations with
quotas being set by a multilaterial management agency through a process
of negotiation. As such, we might reasonably anticipate that for this
species there could be substantial transaction costs involved in
reaching agreement over any quota change, or constraints on the
frequency with which the those changes could occur. Either could
contribute to the observed quota stability.

Such patterns of variable stocks managed under far less variable quotas
are extremely common; for instance, in the management of south Atlantic
snowy grouper (\emph{Epinephelus niveatus}) shown in the righthand
panels (SEDAR 2013). Though the stock assessment shows similarly
frequent variation between 1994 and 2012, the quota has only been
adjusted twice during this period. In contrast to the adjustment in the
bluefin quotas, both adjustments in snowy grouper quotas were phased in
through steps over several years: 1994-96 and 2005-08. More generally,
reviews by Biais (1995) and Patterson (2007) document many cases where
changes in catch quotas that a management agency set were more modest
than changes that would be recommended just by considering variations in
stock abundance.

We use a classic fisheries management question to examine how accounting
for costs of policy adjustment can change optimal policies (see also
Ludwig (1980), Feichtinger et al. (1994), Wirl (1999)). We focus on how
harvest quotas for a stochastically varying fish population can be
chosen to maximize the net present value of a fishery. Our formulation
and solution method largely follow Reed's classic treatment on this
question, a treatment repeated widely in bioeconomic textbooks. We note
that while later work has extended this treatment to deal with a variety
of other issues (e.g. Sethi et al. (2005); Singh et al. (2006); McGough
et al. (2009)), we start from the classical model for simplicity of
presentation and analysis.

With his formulation, Reed (1979) showed that a constant escapement
policy could be optimal under certain conditions. Such a policy involves
choosing annual quotas that are perfectly responsive to recruitment
variation in a fish stock. In poor recruitment years, the quota is set
to zero and no fishing is allowed. Any time a recruitment pulse exceeds
the escapement threshold, a quota is set that allows the fishery to
exactly compensate through harvesting, thereby maintaining the optimal
escapement level. However, in that analysis, Reed did not account for
any costs of policy adjustment, which for such a responsive management
strategy potentially could be large.

@@

\section{Methods}\label{methods}

\subsection{Fish population dynamics (state
equation)}\label{fish-population-dynamics-state-equation}

@@

where \(N_t\) the stock size, \(h_t\) the harvested level, \(Z_t\) gives
the stochastic shocks, which we assume are log-normally distributed and
\(A\) and \(B\) are positive constants.

We assume managers set an annual quota for harvesting \(h_t\) (the
control variable) after observing the stock size that year \(N_t\) (the
state variable), but while still being uncertain about future
environmental conditions and stock sizes. Through time, this gives a
time path of management actions \({\bf h}=(h_1,h_2,\dots)\) that depend
on the stock sizes that were observed (a state dependent control rule).

@@

where \(\mathbf{E}\) is the expectation operator, \(\delta>0\) is the
discount rate and \(\Pi_0\) is the net revenue from operating the
fishery in a given year. In this base case, we assume the annual
dockside (that is, before accounting for any adjustment costs that may
be incurred) net revenue from fishing is

@@

where \(E_t\) represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, \(h_t = q E_t N_t\)
and constant \(q>0\) is the catchability coefficient. In Eqn.
\eqref{eq:annual_netrevenue}, \(p\) is the price per unit harvest and
\(c_0\) the cost of fishing per unit effort and, to simplify the
presentation of results, we assume that these are constants with \(p>0\)
and \(c_0>0\).

To simplify presentation of the results, we will illustrate cases where
the growth parameters are chosen so that the equilibrium biomass for the
equivalent deterministic model without harvesting is 10; (specifically,
\(A=\) 1.5 and \(B=\) 0.05 in Eqn \eqref{eq:state_equation}). To
characterize environmental variability, we assume multiplicative shock
\(Z_t\) is distributed log-normally with log standard deviation
\(\sigma_g\) = 0.2. In addition, we show cases where \(p=\) 10,
\(\delta=\) 0.05 and \(c_0=\) 30. In the supplement we illustrate that
the qualitative patterns observed here are not sensitive to the specific
choices of these parameter values.

Taken together this objective function \eqref{eq:objective} and the
state equation \eqref{eq:state_equation} define a stochastic dynamic
programming problem that we solve using backwards recursion via
Bellman's equation. We denote the resulting state-dependent, optimal
control as \(\mathbf{h}_0^{\ast}\). We solve this problem on a finite
time horizon of \(T =\) 50 using value iteration (Mangel and Clark 1988,
Clark and Mangel 2000).

R code for implementing the dynamic programming algorithm in this
context is provided as a supplementary R package to the paper, along
with scripts for replicating the analyses presented here. To address
concerns about computational reproducibility (Boettiger 2015), a Docker
container image of the software environment is also provided.

\subsection{Costs of policy
adjustment}\label{costs-of-policy-adjustment}

We compare this base case to three alternative problem formulations,
each reflecting different plausible functional forms that costs of
policy adjustment could take. In each, we assume managers can adjust the
quota set in the fishery \(h_t\) in a given year and that any policy
adjustment costs are associated with changes to this control variable.
In each case, we assume there is no cost in initially setting the
harvest policy at time 0.

First we assume that policy adjustment costs are directly proportional
to the magnitude of the change in policy being proposed, such that
larger changes to annual harvesting quotas incur greater policy
adjustment penalties. Specifically, we replace \(\Pi_0\) in the
\(NPV_0\) equation with

@@

where indicator function \(\mathbf{I}(h_t, h_{t-1})\) takes the value 1
if \(h_t=h_{t-1}\) and zero otherwise. For each \(\Pi_i\), we can then
define a new objective function \(NPV_i\) similar to that in Eqn.
\eqref{eq:objective}. We note that the fixed cost has conceptual
analogues to the set-up cost of Reed (1974) and Spulber (1982) but here
the fee is associated with a change in harvesting (\(h_t\neq h_{t-1}\))
and not with harvesting \emph{per se} (\(h_t>0\)) as in those studies.

Taken together with the state equation Eqn. \eqref{eq:state_equation},
each of these new objective functions defines a different stochastic
dynamic programming problem. Again, we solve them numerically using
backwards recursion. To include costs of policy adjustment, we expand
the state space to include both the current stock size and the
management action taken on the previous time step \((N_t,h_{t-1})\). For
each new objective function \(\max NPV_i\), we denote the corresponding
optimal control policy as the vector \(\mathbf{h}_i^*\).

Note that this problem is much larger computationally than the classic
formulation of this stochastic dynamic programming problem for a single
stock. Whereas the classic problem considers each of \(Q\) possible
discrete quotas for \(S\) possible stock values at each time \(t\), for
a search space of size \(S \times Q\) per timestep; this formulation
must also consider all possible values of quota in the \emph{previous}
time step: since the costs that will follow depend on whether and by how
much the harvest policy will change. This creates a total of
\(S \times Q \times Q\) configurations.

In the Supplementary Material, we compare our results with more
conventional fisheries economic formulation in which additional costs
are applied to the control variables themselves, as opposed to
adjustments to the controls:
\(\Pi_4 (N_t,h_t) = p h_t - c_0 E_t-c_4E_t^2\). Additional costs of this
form tend to have a smoothing effect on optimal quotas, but also change
the long-term average stock size or quota size that is optimal.

\subsection{Comparing apples to
apples}\label{comparing-apples-to-apples}

Each policy adjustment cost function is characterized in terms of a cost
coefficient \(c_i\). However, \(c_i\) takes different units for each
functional form. Therefore, if seeking to compare the relative effect of
each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of \(c_i\) so that each has a comparable impact on the optimal
fishery \(NPV\). Fortunately, the optimal control framework provides a
completely natural way to make this comparison: in terms of the
\emph{economic value} (measured by \(NPV\)) of the stock in each case.
We will calibrate our realized economic value under each policy relative
to economic value of the stock when adjustments to the policy are free
(\(NPV_0\)).

Figure 2 shows the calibration graphically. Each curve plots the change
in maximum expected \(NPV\) for a given cost structure as the \(c_i\)
coefficient is increased. In each case, the figure shows maximum
expected NPV with policy adjustment costs as a proportion of the maximum
expected NPV available in the basic problem \(NPV_0({\bf h_0^*})\)
without policy adjustment costs.

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-Figure_2-1.pdf}
\caption{Expected net present value by functional form of penalty.
Horizontal axis shows the coefficient \(c_i\) governing the magnitude of
the policy cost, while vertical axis shows fraction of the maximum
expected net present value dissipated by the policy cost,
\(\tfrac{NPV_0 - NPV_i}{NPV_0}\). The horizontal line indicates a value
of the stock that is reduced by 25\% from the maximum expected value in
the absence of policy adjustment costs, \(NPV_0({\bf h_0^*})\).
Selecting the coefficient \(c_i\) corresponding to this value in each
functional form allows us to make consistent comparisons across the
different functional forms of policy costs.}
\end{figure}

To compare the impact of penalty functions on optimal management across
the different functional forms, we select penalty cost coefficients that
induce the same reduction in maximum expected NPV. For example, the
dashed vertical line in Figure 2 maps the needed cost coefficient
\(c_i\), for each penalty function, such that the fishery is worth 75\%
of its unconstrained value when optimally managed. To demonstrate that
the results shown here are independent of this choice of calibrating to
a 75\% reduction, we conduct the same analysis with a 10\% and 30\%
reduction rate as well. The corresponding figures can be found in the
Supplemental Materials.

\section{Results}\label{results}

\subsection{Effect of policy adjustment costs on optimal quotas and
stock
sizes}\label{effect-of-policy-adjustment-costs-on-optimal-quotas-and-stock-sizes}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-Figure_3-1.pdf}
\caption{Example realization of optimal harvesting strategy and
resulting fish stock sizes under the different functional forms of
adjustment costs. When costs vary linearly with the size of the
adjustment (\(\Pi_1\)), periods with no adjustment are more common than
if adjustments were free. When costs vary quadratically with adjustment
size (\(\Pi_2\)), small adjustments are relatively cheap and thus the
resulting optimal policy always changes, but by a smaller amount than if
adjustments were free. When adjustments incur a fixed fee independent of
size (\(\Pi_3\)), the optimal strategy either remains unchanged or
overshoots the cost-free optimum.}
\end{figure}
